---
title: ""
output: html_document
css: style.css
---

`r Sys.Date()`

### Association between pre-pregnancy BMI and gestational weight gain with size, tempo and velocity of infant weight gain: Analysis of the Newborn Epigenetic Study (NEST) cohort.
  
```{r , results='asis', echo=FALSE, include=FALSE, cache=F}
# library(sitar)
# library(ggplot2)
# library(plyr)
# library(nlme)
# library(doBy)
# library(reshape2)
# library(xtable)
# 
# setwd("~/Dropbox/Projects/baby_growth/data")
# load('m11.Rda')
# load('m12.Rda')
# load('m13.Rda')
# load('m21.Rda')
# load('m22.Rda')
# load('m23.Rda')
# load('nwaz.Rda')
```


#### Table 1 Descriptive Analysis

```{r des, echo=FALSE, results='asis'}
# sex.table <- table(nwaz$sex)
# prop.table(sex.table)
fvar <- subset(nwaz, select=c(mat_gest_wt_gain_c3.f, sex, parity.f, race.f, educ.f, gest_dbt.f, preg_ht.f, smoker.f,bmic5.f))
cvar <- subset(nwaz, select=c(GestAge_TotalDays, mom_age_delv, weight, gest_weight_gain_kg, mat_ht_m))
tblFun <- function(x){
    tbl <- table(x)
    prop <- round(prop.table(tbl)*100,0)
    prop
}
p.table <- apply(fvar, 2, tblFun)
p.table <- data.frame(unlist(p.table))
f.table <- apply(fvar, 2, table)
f.table <- data.frame(unlist(f.table))
n.table <- cbind(f.table, p.table)
for (i in 1:ncol(n.table)) {
  n.table[,i] <- as.character(n.table[,i])
}
# n.table <- data.frame(as.character(n.table[,c(1,2)]))
names(n.table) <- c('N/Mean', 'PCT/SD')

c.m <- data.frame(unlist(apply(cvar, 2, mean, na.rm=T)))
c.sd <- data.frame(unlist(apply(cvar, 2, sd, na.rm=T)))
c.table <- round(cbind(c.m, c.sd), digits=2)
for (i in 1:ncol(c.table)) {
  c.table[,i] <- as.character(c.table[,i])
}
names(c.table) <- c('N/Mean', 'PCT/SD')

des.table <- rbind(n.table, c.table)

des.table <- xtable(des.table, include.rownames=FALSE) # html.table.attributes=list(font.size=7pt)
# digits(des.table)[c(2,6)] <- 0
print(des.table, type='html')

# tblFun <- function(x){
#     tbl <- table(x)
#     tbl <- cbind(tbl,round(prop.table(tbl)*100,0))
#     colnames(tbl) <- c('Count','Percentage')
#     print(names(x))
#     tbl
# }
# tt <- tblFun(fvar$sex)
# do.call(rbind,lapply(fvar,tblFun, use.names=T))

```

#### Table 2 Results of SITAR Models

```{r sitar, results='asis', echo=FALSE}
options(warn=-1,show.error.messages=F)

out.sitar <- function(tmp.mf) {
  tmp.mf.coef <- data.frame(summary(tmp.mf)$tTable)
  tmp.mf.coef$rn <- row.names(tmp.mf.coef)
  tmp.mf.intvl <- data.frame(intervals(tmp.mf)$fixed)
  tmp.mf.intvl$rn <- row.names(tmp.mf.intvl)
  tmp.mf.coef <- merge(tmp.mf.intvl, tmp.mf.coef, by='rn')
  varnames <- colsplit(as.character(tmp.mf.coef$rn), "\\.", c("fe","cov"))
  tmp.mfa.coef <- data.frame(tmp.mf.coef, varnames)
  tmp.mfa.coef <- subset(tmp.mfa.coef, select=c("fe", "cov", "Value", "p.value", "lower", "upper"))
  tmp.mfa.coef <- tmp.mfa.coef[!(tmp.mfa.coef$fe %in% c("s1", "s2", "s3")),]
  tmp.mfa.coef <- tmp.mfa.coef[!(tmp.mfa.coef$cov==""),]
  rownames(tmp.mfa.coef) <- c(1:length(tmp.mfa.coef$fe))
  tmp.mfa <- reshape(data=tmp.mfa.coef, timevar="fe", idvar="cov", direction="wide")
  write.table(tmp.mfa, file="out.csv", sep = ",", col.names = NA, qmethod = "double")
  return(tmp.mfa)
}
m11.out <- out.sitar(m11)
m12.out <- out.sitar(m12)
m13.out <- out.sitar(m13)
m21.out <- out.sitar(m21)
m22.out <- out.sitar(m22)
m23.out <- out.sitar(m23)
names(m11.out) <- c("cov", paste('m11', names(m11.out)[-1], sep=""))
names(m12.out) <- c("cov", paste('m12', names(m12.out)[-1], sep=""))
names(m13.out) <- c("cov", paste('m13', names(m13.out)[-1], sep=""))
names(m21.out) <- c("cov", paste('m21', names(m21.out)[-1], sep=""))
names(m22.out) <- c("cov", paste('m22', names(m22.out)[-1], sep=""))
names(m23.out) <- c("cov", paste('m23', names(m23.out)[-1], sep=""))
all.out <- merge(m11.out, m12.out, by="cov", all=T)
all.out <- merge(all.out, m13.out, by="cov", all=T)
all.out <- merge(all.out, m21.out, by="cov", all=T)
all.out <- merge(all.out, m22.out, by="cov", all=T)
all.out <- merge(all.out, m23.out, by="cov", all=T)
# order.n <- data.frame(cbind(c(21, 18:20, 13, 14, 15:16, 8, 6:7, 9, 17, 1:5, 10, 11:12, 22), 1:22)) # without bmi.c
order.n <- data.frame(cbind(c(22, 19:21, 14, 15, 16:17, 9, 7:8, 10, 18, 1:5, 11, 6, 12:13, 23), 1:23))
order.n <- order.n[order(order.n$X1),]
order.n <- c(order.n$X2)
all.out <- cbind(all.out, order.n)
all.out <- all.out[order(order.n),]
# order.out <- cbind(all.out$cov, 1:length(all.out$cov), order.n)
# format(all.out[,-1], digits=0, nsmall=3, scientific=F) # won't change the actual value
all.out <- cbind(all.out[,1], round(all.out[,-1], digits=3))
# all.out <- data.frame(apply(all.out, 2, as.character)) # already done by paste
all.table <- NULL
all.table <- data.frame(all.out[,1])
names(all.table) <- "Covariate"
# all.table$Model.1a <- paste0(all.out$m11Value.a," (",all.out$m11lower.a,",",all.out$m11upper.a,") ",all.out$m11p.value.a)
p <- function(outd, ind, x) {
  outd <- cbind(outd, paste0(ind[,x+1]," (",ind[,x+3],",",ind[,x+4],") ",ind[,x+2]))
  names(outd)[ncol(outd)] <- paste0("Model.",m,r)
  outd
}
counter <- 1
for (m in 1:6) {
  for (r in c("a", "b", "c")) {
    all.table <- p(all.table, all.out, counter)
    counter <- counter+4
  }
}
all.table[all.table=="NA (NA,NA) NA"] <- NA

# t <- function(d, x, v) {
#   print(head(nwaz[,x+(1:3)]))
#   print(head(paste(d[,x+1], d[,x+2], d[,names(d)==v])))
# }
# t(nwaz, 3, "sex")


all.table <- xtable(all.table, include.rownames=FALSE) # html.table.attributes=list(font.size=7pt)
digits(all.table)[c(2,6)] <- 0
print(all.table, type='html')
```

Note: the values in each cell represent: coefficient (confidence interval) p-value 

#### plot of individual raw data and SITAR adjusted individual data

```{r, echo=FALSE, fig.show='hold'}
plot(m13, opt='u', col=1+as.integer(id) %% 5, ylim=c(2, 18), yfun=exp,
     xlab="Age in Months", ylab="Weight (kg)")

plot(m13, opt='a', col=1+as.integer(id) %% 5, xlim=xaxsd(), ylim=c(2, 18), yfun=exp,
     xlab="Age in Months", ylab="Weight (kg)")
  lines(m13, opt='d', col='red', lwd=2, yfun=exp, xlab="Age in Months", ylab="Weight (kg)")
```

#### plot of growth vs velocity
```{r, echo=FALSE, include=FALSE}
options(warn=-1,show.error.messages=F)

# m23.coef
m23.coef <- data.frame(summary(m23)$tTable)
m23.coef.weight1 <- m23.coef[grep("mat_gest_wt_gain_c3.f1", rownames(m23.coef)), 1]
m23.coef.weight3 <- m23.coef[grep("mat_gest_wt_gain_c3.f3", rownames(m23.coef)), 1]

# m13.coef
m13.coef <- data.frame(summary(m13)$tTable)
m13.coef.bmi6 <- m13.coef[grep("bmic5.f6", rownames(m13.coef)), 1]
```

```{r, echo=FALSE, fig.show='hold'}
plot(m23, y2par=list(col='black'), yfun=exp, xlab="Age in Months", ylab="Weight (kg)", apv=T)
plot(m23, y2par=list(col='black'), yfun=exp, xlab="Age in Months", ylab="Weight (kg)", apv=T, add=T)
plot(m23, y2par=list(col='red'), col='red', yfun=exp, xlab="Age in Months", ylab="Weight (kg)", apv=T,
     add=T, abc=c(a=m23.coef.weight1[1],b=m23.coef.weight1[2],c=m23.coef.weight1[3]))
plot(m23, y2par=list(col='blue'), col='blue', yfun=exp, xlab="Age in Months", ylab="Weight (kg)", apv=T,
     add=T, abc=c(a=m23.coef.weight3[1],b=m23.coef.weight3[2],c=m23.coef.weight3[3]))
legend("right",lty=1,col=c("red", "black","blue"),legend=c("< recommended","= recommanded", "> recommanded"), cex=0.9, inset=0.05)

plot(m13, y2par=list(col='black'), yfun=exp, xlab="Age in Months", ylab="Weight (kg)", apv=T)
plot(m13, y2par=list(col='black'), yfun=exp, xlab="", ylab="", apv=T, add=T)
plot(m13, y2par=list(col='blue'), col='blue', yfun=exp, xlab="", ylab="Weight (kg)", apv=T,
     add=T, abc=c(a=m13.coef.bmi6[1],b=m13.coef.bmi6[2],c=m13.coef.bmi6[3]))
legend("right",lty=1,col=c("black","blue"),legend=c("Normal BMI","BMI > 40"), cex=0.9, inset=0.05)
```
